%% EDBT 2027 regular research-paper submission.
\documentclass[sigconf,review]{acmart}
\geometry{twoside=true,head=13pt,a4paper,includeheadfoot,columnsep=2pc,
  top=57pt,bottom=73pt,inner=54pt,outer=54pt,marginparwidth=2pc,heightrounded}
\usepackage{balance}
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric}
\AtBeginMaketitle{\input{edbt-macros}}
\newcommand{\method}{\textsc{SBC}\xspace}
\newcommand{\yes}{\mathsf{yes}}
\newcommand{\no}{\mathsf{no}}
\begin{document}

\title{Structured Pruning and Calibrated Batched Model Cascades for Hybrid Semantic Queries}
\author{Ann Remizova}
\email{TODO@university.example}
\affiliation{\institution{Univ. Grenoble Alpes, Grenoble INP--Ensimag, LIG}\city{Grenoble}\country{France}}
\renewcommand{\shortauthors}{Remizova}

\begin{abstract}
Queries over heterogeneous data often combine exact predicates on tables with
semantic predicates over text. Executing the latter with a large language model
(LLM) once per surviving tuple is accurate but expensive; batching reduces calls
but can waste inference on tuples that relational operators could eliminate. We
present a structured batched cascade that (i) translates a natural-language query
to a relational filter and semantic predicate, (ii) performs an exact key join,
(iii) classifies candidate pairs in small-model batches, and (iv) coalesces only
uncertain pairs into larger batches for a stronger model. A calibration sample is
labelled by the strong model and selects the least confidence threshold attaining
a target agreement. We implement this design in two existing execution styles:
a row-oriented SUQL executor and a Trummer-style semantic block join. On a
five-query IMDb benchmark with 100 movies, 40 structured candidates, and 12
answers per query, the batched implementation obtains macro precision 0.786,
recall 0.667, and F1 0.701 while averaging 21.84 seconds and 10 model calls. In
the same preliminary run it is 4.28$\times$ faster and uses 13.78$\times$ fewer
calls than the SUQL baseline, while improving macro F1 by 34.8\%. The experiment
has one repetition and synthetic lexical ground truth; we therefore present it
as an implementation study and state the repeated, independently annotated
evaluation required to establish generality.
\end{abstract}
\keywords{semantic query processing, hybrid data, model cascades, batch prompting, query optimization}
\maketitle

\section{Introduction}
A question such as ``Which Horror movies under 100 minutes have reviews
describing them as frightening?'' spans two computational regimes. Genre and
runtime are deterministic relational predicates. Whether prose describes fear
requires semantic interpretation. SUQL makes this composition explicit with
\texttt{answer(text, question)}~\cite{liu2024suql}; a naive physical plan still
invokes the model for many tuples. Semantic block joins amortize prompt overhead
across pairs~\cite{trummer2025semanticjoins}, while model cascades reserve a
larger model for difficult inputs~\cite{chen2023frugalgpt}. These optimizations
act on different terms of execution cost and should be composed.

We study the physical plan
\[
 \text{parse}\rightarrow\text{relational prune}\rightarrow\text{exact join}
 \rightarrow\text{cheap batches}\rightarrow\text{coalesced fallback}.
\]
The central technical issue is routing: a cheap prediction must be accepted only
when its confidence is high enough, yet model scores are query-dependent. Our
implementation samples candidates across the observed confidence range, obtains
strong-model proxy labels, and learns a per-query threshold for a requested
agreement level.

This paper contributes: (1) a database-aware batched cascade for hybrid queries;
(2) a precise account of two current implementations, including their different
confidence semantics; (3) a reproducible four-method IMDb benchmark measuring
quality, latency, and calls; and (4) an analysis of failure behavior and threats
to validity. The result is deliberately not described as a new query language:
it is a physical execution strategy that can sit below one.

\section{Background and Related Work}
SUQL extends SQL with \textsc{Answer} and \textsc{Summary} functions and uses an
LLM semantic parser to translate conversational questions to the hybrid
language~\cite{liu2024suql}. BlendSQL likewise integrates external reasoning
within relational algebra~\cite{glenn2024blendsql}. Our work changes physical
evaluation after parsing rather than expressiveness.

Batch prompting places multiple examples in one request, amortizing fixed prompt
and request costs; its benefit depends on batch size and task complexity
~\cite{cheng2023batch}. Trummer derives block sizes for LLM semantic joins under
context and output constraints~\cite{trummer2025semanticjoins}. We retain exact
key equality outside the model and batch only the residual semantic predicate.

FrugalGPT learns cascades of heterogeneous models to trade quality against cost
~\cite{chen2023frugalgpt}. Our router differs in being local to one query and in
using the strong model as an online proxy oracle. This saves manual labels but
only controls agreement with that model, not correctness against ground truth.

\section{Problem and Cost Model}
Let $M$ be movies, $R$ reviews, $k_M(m)=k_R(r)$ their exact identifier join, and
$q=(p_s,p_u)$ a structured predicate and semantic predicate. The desired movie
set is
\begin{equation}
 A(q)=\{k_M(m):m\in M,\ p_s(m),\ \exists r\in R:\ k_M(m)=k_R(r)\land p_u(r)\}.
\end{equation}
The implementation returns JSON-serializable rows keyed by movie identifier.
Let $N=|M|$, $n=|\sigma_{p_s}(M)\Join_k R|$, cheap and expensive batch sizes be
$b_c,b_e$, and $u(\tau)$ be the number routed to the expensive model at threshold
$\tau$. Ignoring parsing and retries, Trummer V1 makes
\begin{equation}
 C(\tau)=\left\lceil n/b_c\right\rceil+
 \left\lceil m/b_e\right\rceil+\left\lceil u(\tau)/b_e\right\rceil
\end{equation}
calls, where $m\leq\min(n,B)$ is the calibration sample size and $B$ its budget.
Calibration and fallback samples can overlap, so this is a call count, not a
count of unique evaluated pairs. With per-call latency functions $L_c(b)$ and
$L_e(b)$, sequential model time is approximated by the analogous sum. Structured
pruning changes $N$ to $n$; batching changes calls from $O(n)$ to $O(n/b)$; the
cascade changes how many calls use the expensive model.

\section{Execution Strategy}
Figure~\ref{fig:architecture} shows the physical plan. The parser emits a SUQL
query; removing \texttt{answer(...)} leaves SQL executed by SQLite. A conservative
regex extractor is used if parsing fails. Reviews are then restricted by exact
identifier equality, never by an LLM.

\begin{figure*}[t]
\centering
\resizebox{.94\textwidth}{!}{%
\begin{tikzpicture}[node distance=4mm and 5mm,>=Latex,
 every node/.style={font=\footnotesize,inner xsep=3pt,inner ysep=2pt},
 box/.style={draw,rounded corners,align=center,minimum height=7mm,fill=blue!5},
 model/.style={draw,rounded corners,align=center,minimum height=7mm,fill=orange!12},
 gate/.style={draw,diamond,aspect=2.2,align=center,fill=green!10}]
\node[box] (q) {NL question};
\node[model,right=of q] (parse) {SUQL parser};
\node[box,right=of parse] (sql) {SQL pruning\\$\sigma_{p_s}(M)$};
\node[box,right=of sql] (join) {exact ID join};
\node[model,right=of join] (cheap) {cheap batches\\size $b_c$};
\node[gate,right=of cheap] (gate) {$|s_i|\geq\tau$?};
\node[box,above right=4mm and 5mm of gate] (emit) {cheap\\decision};
\node[model,below right=4mm and 5mm of gate] (strong) {strong batches\\size $b_e$};
\node[box,right=13mm of gate] (json) {deduplicate\\and emit JSON};
\draw[->] (q)--(parse);\draw[->] (parse)--(sql);\draw[->] (sql)--(join);
\draw[->] (join)--(cheap);\draw[->] (cheap)--(gate);
\draw[->] (gate)--node[above,sloped]{yes}(emit);\draw[->] (gate)--node[below,sloped]{no}(strong);
\draw[->] (emit.east)-| (json.north);\draw[->] (strong.east)-| (json.south);
\end{tikzpicture}%
}
\caption{Trummer V1 architecture. Calibration (omitted for clarity) samples cheap scores, queries the strong model, and supplies $\tau$ to the gate.}
\label{fig:architecture}
\end{figure*}

\subsection{Two confidence representations}
SUQL V1 scores each row independently. Ideally its score is the binary log odds
\begin{equation} s_i=\log P_c(\yes\mid x_i,q)-\log P_c(\no\mid x_i,q). \end{equation}
When both token log-probabilities are unavailable, the emitted label is assigned
a signed magnitude derived from its probability or a default magnitude 2. Thus
$\hat y_i=\mathbb{1}[s_i\geq0]$ and confidence is $c_i=|s_i|$.

Trummer V1 asks for categorical labels in batches and maps
\begin{equation}
 s_i=\begin{cases}+2&\text{YES}\cr 0&\text{UNCERTAIN}\cr-2&\text{NO.}\end{cases}
\end{equation}
Consequently its ``confidence'' currently has only two magnitudes. This is a
routing heuristic, not a calibrated probability; a threshold learned from these
scores can only separate UNCERTAIN from hard decisions (or disable all cheap
decisions). The distinction is important when interpreting calibration results.

\subsection{Online agreement threshold}
Let $S$ be at most $B$ scored candidates, selected approximately uniformly by
rank over descending confidence. The expensive model supplies proxy labels
$y_i^e$. For each observed candidate threshold $t$, define
\begin{equation}
 \widehat a(t)=\frac{\sum_{i\in S}\mathbf1[c_i\geq t]\mathbf1[\hat y_i=y_i^e]}
 {\sum_{i\in S}\mathbf1[c_i\geq t]}.
\end{equation}
For target agreement $\alpha=0.9$, the implementation scans distinct confidence
values in increasing order and returns the first $t$ with $\widehat a(t)\geq
\alpha$. If none exists, $\tau=\bot$ and every scored candidate is escalated.
With a manual threshold, calibration is skipped. Missing cheap scores also route
to fallback. A cheap score is accepted iff $c_i\geq\tau$; its sign determines
accept or reject.

\begin{figure}[t]
\small
\begin{tabular}{p{.97\columnwidth}}
\toprule
\textbf{Algorithm 1: structured batched cascade}\\\midrule
1. Parse $q$ into $(p_s,p_u)$; on failure extract only safe predicates.\\
2. $M'\leftarrow\sigma_{p_s}(M)$; $X\leftarrow M'\Join_{k_M=k_R}R$.\\
3. For each block $G\subseteq X$, $|G|\le b_c$, obtain cheap scores $s_i$.\\
4. Sample $S\subseteq X$, $|S|\le B$ across confidence ranks; obtain strong labels and learn $\tau$.\\
5. Accept confident positive scores; reject confident negative scores; append all other candidates to $U$.\\
6. For blocks $H\subseteq U$, $|H|\le b_e$, obtain strong labels and accept positives.\\
7. Deduplicate accepted rows by movie ID and return them.\\\bottomrule
\end{tabular}
\end{figure}

\subsection{Prompt and failure contracts}
Both prompts demand one identifier-bound decision per candidate, temperature zero,
and no prose. The cheap Trummer prompt defines YES/UNCERTAIN/NO; the expensive
prompt defines recall-first YES/NO. Reviews are truncated to 3,500 characters;
the SUQL scorer uses 1,800 and its expensive fallback 6,000. A malformed or
incomplete cheap batch escalates its candidates. A failed Trummer expensive batch
is conservatively rejected by producing no accepted IDs; SUQL retries transport
errors and an invalid answer, then returns NO. These fail-closed choices improve
precision but can reduce recall.

\section{Experimental Methodology}
The canonical five-query suite covers recommendation/year, fear/genre and
runtime, visual praise/genre, pacing/director, and originality/title. Each query
contains 100 movies, 40 structured candidates, and 12 answer IDs. Labels are
generated by deterministic lexical patterns and include evidence and rationales;
no evaluated model creates ground truth. We compare SUQL baseline, SUQL V1,
Trummer adaptive block join, and Trummer V1. The reported run uses Gemma~4 e2b
as cheap model and Gemma~4 26b as expensive model, temperature zero, $b_c=8$,
$b_e=32$, $B=20$, and $\alpha=.9$. Metrics are macro-averaged across questions:
$P=TP/(TP+FP)$, $R=TP/(TP+FN)$, and $F_1=2PR/(P+R)$.

\section{Results}
\begin{table}[t]
\caption{Five-query preliminary run (one repetition). Calls and seconds are means per query; quality is macro-averaged.}
\label{tab:results}\centering\small
\begin{tabular}{lrrrrr}\toprule
Method&P&R&F1&sec.&calls\\\midrule
SUQL baseline&.537&.567&.520&93.39&137.8\\
SUQL V1&.617&.717&.638&419.12&68.0\\
Trummer baseline&.606&.333&.396&40.41&100.4\\
Trummer V1&\textbf{.786}&.667&\textbf{.701}&\textbf{21.84}&\textbf{10.0}\\\bottomrule
\end{tabular}
\end{table}
Trummer V1 reduces latency by 76.6\% and calls by 92.7\% relative to SUQL,
while raising F1 by 34.8\%. Relative to the Trummer baseline, it reduces latency
46.0\% and calls 90.0\%, and raises F1 77.0\%. It uses five cheap and five
expensive calls per question; calibration calls are included. SUQL V1 improves
quality but is 4.49$\times$ slower than its baseline because it scores 40 rows
individually and performs row-wise calibration/fallback. Thus cascading alone
does not guarantee lower latency: batching and the fraction escalated determine
the outcome.

\section{Limitations and Future Work}
The primary result has one repetition, one domain, five predicates, locally
served models, and lexically constructed labels. It does not support confidence
intervals or claims about arbitrary semantic queries. Strong-model agreement is
not correctness, and calibration reuses the evaluated query population. Calls
ignore prompt/completion tokens and model sizes. The four methods also implement
different physical operators, so the comparison is end-to-end rather than a
factorial ablation.

The next evaluation should use independently annotated data, at least ten repeated
runs with fixed software/model versions, per-query distributions and paired
bootstrap intervals, token and energy measurements, and ablations for pruning,
batching, calibration, and cascade. Replacing categorical Trummer scores with
log-odds or calibrated probabilities would make $\tau$ genuinely granular.

\section{Conclusion}
Structured pruning, batching, and heterogeneous model routing remove three
different sources of waste. Their composition yields the strongest preliminary
quality/cost point in our benchmark, whereas a row-wise cascade can be slower
despite fewer expensive calls. The mathematical account also exposes a current
limitation: categorical batch decisions provide only coarse confidence. The
promising result therefore motivates, rather than substitutes for, a repeated
and independently annotated study.

\section{Artifacts}
The supplement contains all four implementations, exact prompts, data provenance,
annotations, manifests, runners, raw aggregate tables, and plotting scripts.

\balance
\bibliographystyle{ACM-Reference-Format}
\bibliography{references}
\end{document}
